---
title: "VAR vs Local Projections: a simple comparison with Brazilian data"
author: "Ana Dutra
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Here I use two methods — VAR (Vector Autoregression) and LPs (Local Projections) — to compare the results of this simple analysis.

First, we need to understand what a VAR is:
- It is a system of equations where each variable depends on its own lags and the lags of the others.
- There is no economic theory behind it yet: it only captures the raw dynamics of the data.
- We can apply Cholesky decomposition to identify structural shocks.

And we also need to understand Local Projections:
- With LPs we obtain impulse responses directly, unlike VARs where we first estimate the model and then derive the responses.
- We can estimate the impulse response for each horizon \(h\) separately.
- The shocks can come from an identified VAR or from an external instrument.
- LPs are more robust to specification errors and more flexible.

## Installing packages

```{r packages, eval=FALSE}
install.packages(c("rbcb", "vars", "lpirfs", "dplyr", "tidyr", "purrr", "ggplot2"))
```

## Getting the data

```{r data}
library(rbcb)
library(tidyr)
library(dplyr)
library(purrr)

ibcbr <- get_series(24364, start_date = "2003-01-01")
ipca  <- get_series(433,   start_date = "2003-01-01")
selic <- get_series(4189,  start_date = "2003-01-01")

dados <- reduce(list(ibcbr, ipca, selic), full_join, by = "date") |>
  rename(pib = "24364", infla = "433", juros = "4189") |>
  drop_na()
```

The Selic rate is expressed in annual terms (% p.a.). The ideal procedure would be to convert it to an effective monthly rate (divide by 100 and then compute \((1+\text{rate})^{1/12} - 1\)), but for this simple exercise we keep the annualized rate because we are interested only in the direction of the shock.

## Transformations to guarantee stationarity

```{r transform}
dados <- dados |>
  mutate(
    ln_pib =log(pib),
    d_infla =infla,        # IPCA is already a monthly change
    juros_nivel =juros        # level (annualised)
  )
```

## Estimate the VAR in reduced form

```{r var}
library(vars)

data_var <- dados |>
  dplyr::select(ln_pib, d_infla, juros_nivel) |>
  na.omit()

# Choose the lag length by information criteria
lag_selection <- VARselect(data_var, lag.max = 13, type = "const")
lag_selection$selection

p <- 3  # chosen lag order

var_model <- VAR(data_var, p = p, type = "const")
summary(var_model)
```

## Impulse-response function via Cholesky

### (Order: PIB → Inflation → Interest rate)

```{r irf_var}
# Order: 1 = PIB, 2 = Inflation, 3 = Interest rate (the last receives the monetary policy shock)
irf_var <- irf(var_model, impulse = "juros_nivel", response = c("ln_pib", "d_infla"),
               n.ahead = 24, ortho = TRUE, cumulative = FALSE, boot = TRUE)

plot(irf_var)
```

The "price puzzle" appears — inflation rises before it falls, a known problem that Cholesky identification often does not solve.

## Local Projections (Jordà approach)

We use the VAR residuals as a "monetary policy shock" (identified by Cholesky). This is just to compare the two methodologies using the same shock, though it is not the ideal procedure.

```{r lp}
library(lpirfs)

resultados_lp <- lp_lin(data_var,
                        lags_endog_lin = p,
                        trend = 0,
                        shock_type = 1,
                        confint = 1.96,
                        hor = 24)

plot(resultados_lp)
```

For the interest rate, the confidence bands are wider. LPs are more robust but less efficient.